<!---
  Licensed to the Apache Software Foundation (ASF) under one
  or more contributor license agreements.  See the NOTICE file
  distributed with this work for additional information
  regarding copyright ownership.  The ASF licenses this file
  to you under the Apache License, Version 2.0 (the
  "License"); you may not use this file except in compliance
  with the License.  You may obtain a copy of the License at

    http://www.apache.org/licenses/LICENSE-2.0

  Unless required by applicable law or agreed to in writing,
  software distributed under the License is distributed on an
  "AS IS" BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
  KIND, either express or implied.  See the License for the
  specific language governing permissions and limitations
  under the License.
-->

# Manipulating Data - Arrays

## Introduction

An Arrow Array is roughly equivalent to an R vector - it can be used to 
represent a single column of data, with all values having the same data type.  

A number of base R functions which have S3 generic methods have been implemented 
to work on Arrow Arrays; for example `mean`, `min`, and `max`.  

## Filter by values matching a predicate or mask

You want to search for values in an Array that match a predicate condition.

### Solution

```{r, array_predicate}
my_values <- Array$create(c(1:5, NA))
my_values[my_values > 3]
```
```{r, test_array_predicate, opts.label = "test"}
test_that("array_predicate works as expected", {
  expect_equal(my_values[my_values > 3], Array$create(c(4L, 5L, NA)))
})
```

### Discussion

You can refer to items in an Array using the square brackets `[]` like you can
an R vector.

## Compute Mean/Min/Max, etc value of an Array

You want to calculate the mean, minimum, or maximum of values in an array.

### Solution

```{r, array_mean_na}
my_values <- Array$create(c(1:5, NA))
mean(my_values, na.rm = TRUE)
```
```{r, test_array_mean_na, opts.label = "test"}
test_that("array_mean_na works as expected", {
  expect_equal(mean(my_values, na.rm = TRUE), Scalar$create(3))
})
```

### Discussion

Many base R generic functions such as `mean()`, `min()`, and `max()` have been
mapped to their Arrow equivalents, and so can be called on Arrow Array objects 
in the same way. They will return Arrow objects themselves.

If you want to use an R function which does not have an Arrow mapping, you can 
use `as.vector()` to convert Arrow objects to base R vectors.

```{r, array_fivenum}
arrow_array <- Array$create(1:100)
# get Tukey's five-number summary
fivenum(as.vector(arrow_array))
```
```{r, test_array_fivenum, opts.label = "test"}

test_that("array_fivenum works as expected", {
  
  expect_identical(
    fivenum(as.vector(arrow_array)),
    c(1, 25.5, 50.5, 75.5, 100)
  )
})

```

You can tell if a function is a standard S3 generic function by looking 
at the body of the function - S3 generic functions call `UseMethod()`
to determine the appropriate version of that function to use for the object.

```{r}
mean
```

You can also use `isS3stdGeneric()` to determine if a function is an S3 generic.

```{r}
isS3stdGeneric("mean")
```

If you find an S3 generic function which isn't implemented for Arrow objects 
but you would like to be able to use, please 
[open an issue on the project JIRA](https://issues.apache.org/jira/projects/ARROW/issues).

## Count occurrences of elements in an Array

You want to count repeated values in an Array.

### Solution

```{r, value_counts}
repeated_vals <- Array$create(c(1, 1, 2, 3, 3, 3, 3, 3))
value_counts(repeated_vals)
```

```{r, test_value_counts, opts.label = "test"}
test_that("value_counts works as expected", {
  expect_equal(
    as.vector(value_counts(repeated_vals)),
    tibble(
      values = as.numeric(names(table(as.vector(repeated_vals)))),
      counts = as.vector(table(as.vector(repeated_vals)))
    )
  )
})
```

### Discussion

Some functions in the Arrow R package do not have base R equivalents. In other 
cases, the base R equivalents are not generic functions so they cannot be called
directly on Arrow Array objects.

For example, the `value_counts()` function in the Arrow R package is loosely 
equivalent to the base R function `table()`, which is not a generic function. 

## Apply arithmetic functions to Arrays.

You want to use the various arithmetic operators on Array objects.

### Solution

```{r, add_array}
num_array <- Array$create(1:10)
num_array + 10
```
```{r, test_add_array, opts.label = "test"}
test_that("add_array works as expected", {
  # need to specify expected array as 1:10 + 10 instead of 11:20 so is double not integer
  expect_equal(num_array + 10, Array$create(1:10 + 10))
})
```

### Discussion

You will get the same result if you pass in the value you're adding as an Arrow object.

```{r, add_array_scalar}
num_array + Scalar$create(10)
```
```{r, test_add_array_scalar, opts.label = "test"}
test_that("add_array_scalar works as expected", {
  # need to specify expected array as 1:10 + 10 instead of 11:20 so is double not integer
  expect_equal(num_array + Scalar$create(10), Array$create(1:10 + 10))
})
```

## Call Arrow compute functions directly on Arrays

You want to call an Arrow compute function directly on an Array.

### Solution

```{r, call_function}
first_100_numbers <- Array$create(1:100)

# Calculate the variance of 1 to 100, setting the delta degrees of freedom to 0.
call_function("variance", first_100_numbers, options = list(ddof = 0))

```
```{r, test_call_function, opts.label = "test"}
test_that("call_function works as expected", {
  expect_equal(
    call_function("variance", first_100_numbers, options = list(ddof = 0)),
    Scalar$create(833.25)
  )
})
```
### Discussion

You can use `call_function()` to call Arrow compute functions directly on 
Scalar, Array, and ChunkedArray objects.  The returned object will be an Arrow object.

### See also

For a more in-depth discussion of Arrow compute functions, see the section on [using arrow functions in dplyr verbs in arrow](#use-arrow-functions-in-dplyr-verbs-in-arrow)
