# Creating Arrow Objects

## Build an Arrow Table from native language types

### Manually create a Table from an R object

You may want to convert an existing data frame in R to an Arrow Table object.

```{r, table_create}
# Create an example data frame
my_tibble <- tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99))
# Convert to Arrow Table
my_table <- Table$create(my_tibble)
# View table
my_table
```
```{r, test_table_create, opts.label = "test"}
test_that("table_create works as expected", {
  expect_s3_class(my_table, "Table")
  expect_identical(dplyr::collect(my_table), my_tibble)
})
```
#### View the contents of an Arrow Table

You can view the contents of an Arrow Table using `dplyr::collect()`

```{r, table_collect}
# View Table
dplyr::collect(my_table)
```
```{r, test_table_collect, opts.label = "test"}
test_that("table_collect works as expected", {
  expect_identical(dplyr::collect(my_table), my_tibble)
})
```

### Manually create a RecordBatch

You may want to convert an existing data frame in R to an Arrow RecordBatch object.

```{r, record_batch_create}
# Create an example data frame
my_tibble <- tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99))
# Convert to Arrow RecordBatch
my_record_batch <- record_batch(my_tibble)
# View RecordBatch
my_record_batch
```
```{r, test_record_batch_create, opts.label = "test"}
test_that("record_batch_create works as expected", {
  expect_s3_class(my_record_batch, "RecordBatch")
  expect_identical(dplyr::collect(my_record_batch), my_tibble)
})
```
#### View the contents of a RecordBatch

You can view the contents of a RecordBatch using `dplyr::collect()`

```{r, rb_collect}
# View RecordBatch
dplyr::collect(my_record_batch)
```
```{r, test_rb_collect, opts.label = "test"}
test_that("rb_collect works as expected", {
  expect_identical(dplyr::collect(my_record_batch), my_tibble)
})
```

## Storing Categorical Data in Arrow

An Arrow Dictionary object is similar to a factor in R, in that it allows for efficient storage of categorical data by allowing you to map between indices and values, reducing the amount of storage.  If you have an R data frame containing factors, converting it to an Arrow object will automatically encode that column as a dictionary.

```{r}
class(iris$Species)
```

```{r, dictionary}
iris_rb <- record_batch(iris)
iris_rb
```
```{r, test_dictionary, opts.label = "test"}
test_that("dictionary works as expected", {
  expect_s3_class(iris_rb$Species, "DictionaryArray")
})
```
