# Creating Arrow Objects

## Create an Arrow Array from an R object

You to convert an existing vector in R to an Arrow Array object.

### Solution

```{r, array_create}
# Create an example vector
score = c(99, 97, 99)

# Convert to Arrow Array
score_array <- Array$create(score)

# View Array
score_array
```
```{r, test_array_create, opts.label = "test"}
test_that("array_create works as expected", {
  expect_s3_class(score_array, "Array")
  expect_identical(as.vector(score_array), score)
})
```

## Create a Arrow Table from an R object

You want to convert an existing data frame in R to an Arrow Table object.

### Solution

```{r, table_create}
# Create an example data frame
my_tibble <- tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99))
# Convert to Arrow Table
my_table <- Table$create(my_tibble)
# View table
my_table
```
```{r, test_table_create, opts.label = "test"}
test_that("table_create works as expected", {
  expect_s3_class(my_table, "Table")
  expect_identical(dplyr::collect(my_table), my_tibble)
})
```
## View the contents of an Arrow Table or RecordBatch

You want to view the contents of an Arrow Table or RecordBatch.

### Solution

```{r, table_collect}
# View Table
dplyr::collect(my_table)
```
```{r, test_table_collect, opts.label = "test"}
test_that("table_collect works as expected", {
  expect_identical(dplyr::collect(my_table), my_tibble)
})
```

## Manually create a RecordBatch from an R object.

You want to convert an existing data frame in R to an Arrow RecordBatch object.

### Solution

```{r, record_batch_create}
# Create an example data frame
my_tibble <- tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99))
# Convert to Arrow RecordBatch
my_record_batch <- record_batch(my_tibble)
# View RecordBatch
my_record_batch
```
```{r, test_record_batch_create, opts.label = "test"}
test_that("record_batch_create works as expected", {
  expect_s3_class(my_record_batch, "RecordBatch")
  expect_identical(dplyr::collect(my_record_batch), my_tibble)
})
```
