# Creating Arrow Objects

## Manually create an Array from an R object

You to convert an existing vector in R to an Arrow Array object.

### Solution

```{r, array_create}
# Create an example vector
score = c(99, 97, 99)

# Convert to Arrow Array
score_array <- Array$create(score)

# View Array
score_array
```
```{r, test_array_create, opts.label = "test"}
test_that("array_create works as expected", {
  expect_s3_class(score_array, "Array")
  expect_identical(as.vector(score_array), score)
})
```

## Manually create a Table from an R object

You want to convert an existing data frame in R to an Arrow Table object.

### Solution

```{r, table_create}
# Create an example data frame
my_tibble <- tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99))
# Convert to Arrow Table
my_table <- Table$create(my_tibble)
# View table
my_table
```
```{r, test_table_create, opts.label = "test"}
test_that("table_create works as expected", {
  expect_s3_class(my_table, "Table")
  expect_identical(dplyr::collect(my_table), my_tibble)
})
```
## View the contents of an Arrow Table or RecordBatch

You want to view the contents of an Arrow Table or RecordBatch.

### Solution

```{r, table_collect}
# View Table
dplyr::collect(my_table)
```
```{r, test_table_collect, opts.label = "test"}
test_that("table_collect works as expected", {
  expect_identical(dplyr::collect(my_table), my_tibble)
})
```

## Manually create a RecordBatch from an R object.

You want to convert an existing data frame in R to an Arrow RecordBatch object.

### Solution

```{r, record_batch_create}
# Create an example data frame
my_tibble <- tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99))
# Convert to Arrow RecordBatch
my_record_batch <- record_batch(my_tibble)
# View RecordBatch
my_record_batch
```
```{r, test_record_batch_create, opts.label = "test"}
test_that("record_batch_create works as expected", {
  expect_s3_class(my_record_batch, "RecordBatch")
  expect_identical(dplyr::collect(my_record_batch), my_tibble)
})
```

## Storing Categorical Data in Arrow

<!-- Note: this entire section should be moved to the "working with data types" section once that PR has been merged -->

You want to store categorical data in Arrow.

### Solution

```{r}
class(iris$Species)
```

```{r, dictionary}
iris_rb <- record_batch(iris)
iris_rb
```
```{r, test_dictionary, opts.label = "test"}
test_that("dictionary works as expected", {
  expect_s3_class(iris_rb$Species, "DictionaryArray")
})
```

### Discussion

An Arrow Dictionary object is similar to a factor in R, in that it allows for 
efficient storage of categorical data by allowing you to map between indices 
and values, reducing the amount of storage.  If you have an R data frame 
containing factors, converting it to an Arrow object will automatically encode 
that column as a dictionary.

