<!---
  Licensed to the Apache Software Foundation (ASF) under one
  or more contributor license agreements.  See the NOTICE file
  distributed with this work for additional information
  regarding copyright ownership.  The ASF licenses this file
  to you under the Apache License, Version 2.0 (the
  "License"); you may not use this file except in compliance
  with the License.  You may obtain a copy of the License at

    http://www.apache.org/licenses/LICENSE-2.0

  Unless required by applicable law or agreed to in writing,
  software distributed under the License is distributed on an
  "AS IS" BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
  KIND, either express or implied.  See the License for the
  specific language governing permissions and limitations
  under the License.
-->

# Creating Arrow Objects {#creating-arrow-objects}

## Create an Arrow Array from an R object

You want to convert an existing vector in R to an Arrow Array object.

### Solution

```{r, array_create}
# Create an example vector
score = c(99, 97, 99)

# Convert to Arrow Array
score_array <- Array$create(score)

# View Array
score_array
```
```{r, test_array_create, opts.label = "test"}
test_that("array_create works as expected", {
  expect_s3_class(score_array, "Array")
  expect_identical(as.vector(score_array), score)
})
```

## Create a Arrow Table from an R object

You want to convert an existing data frame in R to an Arrow Table object.

### Solution

```{r, table_create}
# Create an example data frame
my_tibble <- tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99))
# Convert to Arrow Table
my_table <- arrow_table(my_tibble)
# View table
my_table
```
```{r, test_table_create, opts.label = "test"}
test_that("table_create works as expected", {
  expect_s3_class(my_table, "Table")
  expect_identical(dplyr::collect(my_table), my_tibble)
})
```
## View the contents of an Arrow Table or RecordBatch

You want to view the contents of an Arrow Table or RecordBatch.

### Solution

```{r, table_collect}
# View Table
dplyr::collect(my_table)
```
```{r, test_table_collect, opts.label = "test"}
test_that("table_collect works as expected", {
  expect_identical(dplyr::collect(my_table), my_tibble)
})
```

## Manually create a RecordBatch from an R object.

You want to convert an existing data frame in R to an Arrow RecordBatch object.

### Solution

```{r, record_batch_create}
# Create an example data frame
my_tibble <- tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99))
# Convert to Arrow RecordBatch
my_record_batch <- record_batch(my_tibble)
# View RecordBatch
my_record_batch
```
```{r, test_record_batch_create, opts.label = "test"}
test_that("record_batch_create works as expected", {
  expect_s3_class(my_record_batch, "RecordBatch")
  expect_identical(dplyr::collect(my_record_batch), my_tibble)
})
```
