# Flight

## Introduction

Flight is a general-purpose client-server framework for high performance 
transport of large datasets over network interfaces, built as part of the 
Apache Arrow project.

Flight allows for highly efficient data transfer as it:

* removes the need for deserialization during data transfer
* allows for parallel data streaming
* is highly optimized to take advantage of Arrow’s columnar format.

The arrow package provides methods for connecting to Flight RPC servers to send 
and receive data.

## Connect to a Flight server

You want to connect to a Flight server running on a specified host and port.

```{r, eval = FALSE}
local_client <- flight_connect(host = "127.0.0.1", port = 8089)
```

### Discussion

For an example of how to set up a Flight server from R, see 
[the Flight vignette](https://arrow.apache.org/docs/r/articles/flight.html).

## Send data to a Flight server

You want to send data that you have in memory to a Flight server

```{r, eval = FALSE}
# Connect to the Flight server
local_client <- flight_connect(host = "127.0.0.1", port = 8089)

# Send the data
flight_put(
  local_client,
  data = airquality,
  path = "pollution_data"
)
```

## Check what resources exist on a Flight server

You want to see what paths are available on a Flight server.

```{r, eval = FALSE}
list_flights(local_client)
```

```{r}
# [1] "pollution_data"
```


## Retrieve data from a Flight server

You want to retrieve data on a Flight server from a specified path.

```{r, eval = FALSE}
flight_get(
  local_client,
  "pollution_data"
)
```

```{r, eval = FALSE}
# Table
# 153 rows x 6 columns
# $Ozone <int32>
# $Solar.R <int32>
# $Wind <double>
# $Temp <int32>
# $Month <int32>
# $Day <int32>
# 
# See $metadata for additional Schema metadata
```
