<!---
  Licensed to the Apache Software Foundation (ASF) under one
  or more contributor license agreements.  See the NOTICE file
  distributed with this work for additional information
  regarding copyright ownership.  The ASF licenses this file
  to you under the Apache License, Version 2.0 (the
  "License"); you may not use this file except in compliance
  with the License.  You may obtain a copy of the License at

    http://www.apache.org/licenses/LICENSE-2.0

  Unless required by applicable law or agreed to in writing,
  software distributed under the License is distributed on an
  "AS IS" BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
  KIND, either express or implied.  See the License for the
  specific language governing permissions and limitations
  under the License.
--> 

# Flight

## Introduction

Flight is a general-purpose client-server framework for high performance 
transport of large datasets over network interfaces, built as part of the 
Apache Arrow project.

Flight allows for highly efficient data transfer as it:

* removes the need for serialization during data transfer
* allows for parallel data streaming
* is highly optimized to take advantage of Arrow’s columnar format.

The arrow package provides methods for connecting to Flight RPC servers to send 
and receive data.

It should be noted that the Flight implementation in the R package depends on 
[PyArrow](https://arrow.apache.org/docs/python/) which is called via 
[reticulate](https://rstudio.github.io/reticulate/). This is quite different 
from the other capabilities in the R package, nearly all of which are all 
implemented directly.

## Connect to a Flight server

You want to connect to a Flight server running on a specified host and port.

### Solution

```{r, eval = FALSE}
local_client <- flight_connect(host = "127.0.0.1", port = 8089)
```

### See also

For an example of how to set up a Flight server from R, see 
[the Flight vignette](https://arrow.apache.org/docs/r/articles/flight.html).

## Send data to a Flight server

You want to send data that you have in memory to a Flight server

### Solution

```{r, eval = FALSE}
# Connect to the Flight server
local_client <- flight_connect(host = "127.0.0.1", port = 8089)

# Send the data
flight_put(
  local_client,
  data = airquality,
  path = "pollution_data"
)
```

## Check what resources exist on a Flight server

You want to see what paths are available on a Flight server.

### Solution

```{r, eval = FALSE}
# Connect to the Flight server
local_client <- flight_connect(host = "127.0.0.1", port = 8089)

# Retrieve path listing
list_flights(local_client)
```

```{r}
# [1] "pollution_data"
```


## Retrieve data from a Flight server

You want to retrieve data on a Flight server from a specified path.

### Solution

```{r, eval = FALSE}
# Connect to the Flight server
local_client <- flight_connect(host = "127.0.0.1", port = 8089)

# Retrieve data
flight_get(
  local_client,
  "pollution_data"
)
```

```{r, eval = FALSE}
# Table
# 153 rows x 6 columns
# $Ozone <int32>
# $Solar.R <int32>
# $Wind <double>
# $Temp <int32>
# $Month <int32>
# $Day <int32>
# 
# See $metadata for additional Schema metadata
```
