# Manipulating Data

## Computing Mean/Min/Max, etc value of an Array

Many base R generic functions such as `mean()`, `min()`, and `max()` have been mapped to their Arrow equivalents, and so can be called on Arrow Array objects in the same way. They will return Arrow objects themselves.

```{r, array_mean_na}
my_values <- Array$create(c(1:5, NA))
mean(my_values, na.rm = TRUE)
```
```{r, test_array_mean_na, opts.label = "test"}
test_that("array_mean_na works as expected", {
  expect_equal(mean(my_values, na.rm = TRUE), Scalar$create(3))
})
```
If you want to use an R function which does not have an Arrow mapping, you can use `as.vector()` to convert Arrow objects to base R vectors.

```{r, fivenum}
fivenum(as.vector(my_values))
```
```{r, test_fivenum, opts.label = "test"}
test_that("fivenum works as expected", {
  expect_equal(fivenum(as.vector(my_values)), 1:5)
})
```

## Counting occurrences of elements in an Array

Some functions in the Arrow R package do not have base R equivalents. In other cases, the base R equivalents are not generic functions so they cannot be called directly on Arrow Array objects.

For example, the `value_count()` function in the Arrow R package is loosely equivalent to the base R function `table()`, which is not a generic function. To count the elements in an R vector, you can use `table()`; to count the elements in an Arrow Array, you can use `value_count()`.

```{r, value_counts}
repeated_vals <- Array$create(c(1, 1, 2, 3, 3, 3, 3, 3))
value_counts(repeated_vals)
```

```{r, test_value_counts, opts.label = "test"}
test_that("value_counts works as expected", {
  expect_equal(
    as.vector(value_counts(repeated_vals)),
    tibble(
      values = as.numeric(names(table(as.vector(repeated_vals)))),
      counts = as.vector(table(as.vector(repeated_vals)))
    )
  )
})
```

## Applying arithmetic functions to Arrays.

You can use the various arithmetic operators on Array objects.

```{r, add_array}
num_array <- Array$create(1:10)
num_array + 10
```
```{r, test_add_array, opts.label = "test"}
test_that("add_array works as expected", {
  # need to specify expected array as 1:10 + 10 instead of 11:20 so is double not integer
  expect_equal(num_array + 10, Array$create(1:10 + 10))
})
```

You will get the same result if you pass in the value you're adding as an Arrow object.

```{r, add_array_scalar}
num_array + Scalar$create(10)
```
```{r, test_add_array_scalar, opts.label = "test"}
test_that("add_array_scalar works as expected", {
  # need to specify expected array as 1:10 + 10 instead of 11:20 so is double not integer
  expect_equal(num_array + Scalar$create(10), Array$create(1:10 + 10))
})
```
