# Manipulating Data - Arrays

## Computing Mean/Min/Max, etc value of an Array

You want to calculate the mean, minimum, or maximum of values in an array.

### Solution

```{r, array_mean_na}
my_values <- Array$create(c(1:5, NA))
mean(my_values, na.rm = TRUE)
```
```{r, test_array_mean_na, opts.label = "test"}
test_that("array_mean_na works as expected", {
  expect_equal(mean(my_values, na.rm = TRUE), Scalar$create(3))
})
```

### Discussion

Many base R generic functions such as `mean()`, `min()`, and `max()` have been
mapped to their Arrow equivalents, and so can be called on Arrow Array objects 
in the same way. They will return Arrow objects themselves.

If you want to use an R function which does not have an Arrow mapping, you can 
use `as.vector()` to convert Arrow objects to base R vectors.

```{r, fivenum}
fivenum(as.vector(my_values))
```
```{r, test_fivenum, opts.label = "test"}
test_that("fivenum works as expected", {
  expect_equal(fivenum(as.vector(my_values)), 1:5)
})
```

## Counting occurrences of elements in an Array

You want to count repeated values in an Array.

### Solution

```{r, value_counts}
repeated_vals <- Array$create(c(1, 1, 2, 3, 3, 3, 3, 3))
value_counts(repeated_vals)
```

```{r, test_value_counts, opts.label = "test"}
test_that("value_counts works as expected", {
  expect_equal(
    as.vector(value_counts(repeated_vals)),
    tibble(
      values = as.numeric(names(table(as.vector(repeated_vals)))),
      counts = as.vector(table(as.vector(repeated_vals)))
    )
  )
})
```

### Discussion

Some functions in the Arrow R package do not have base R equivalents. In other 
cases, the base R equivalents are not generic functions so they cannot be called
directly on Arrow Array objects.

For example, the `value_count()` function in the Arrow R package is loosely 
equivalent to the base R function `table()`, which is not a generic function. 

## Applying arithmetic functions to Arrays.

You want to use the various arithmetic operators on Array objects.

### Solution

```{r, add_array}
num_array <- Array$create(1:10)
num_array + 10
```
```{r, test_add_array, opts.label = "test"}
test_that("add_array works as expected", {
  # need to specify expected array as 1:10 + 10 instead of 11:20 so is double not integer
  expect_equal(num_array + 10, Array$create(1:10 + 10))
})
```

### Discussion

You will get the same result if you pass in the value you're adding as an Arrow object.

```{r, add_array_scalar}
num_array + Scalar$create(10)
```
```{r, test_add_array_scalar, opts.label = "test"}
test_that("add_array_scalar works as expected", {
  # need to specify expected array as 1:10 + 10 instead of 11:20 so is double not integer
  expect_equal(num_array + Scalar$create(10), Array$create(1:10 + 10))
})
```

## Using base R or tidyverse functions with arrow Arrays
