# Reading and Writing Data

This chapter contains recipes related to reading and writing data using Apache Arrow.  When reading data using Apache Arrow, there are 2 different ways you may choose to read in the data:
1. a `tibble`
2. an Arrow Table

There are a number of circumstances in which you may want to read in the data as an Arrow Table:
* your dataset is large and if you load it into memory, it may lead to performance issues
* you want faster performance from your `dplyr` queries
* you want to be able to take advantage of Arrow's compute functions

## Converting from a tibble to an Arrow Table

You can convert an existing `tibble` or `data.frame` into an Arrow Table.

```{r, table_create_from_tibble}
air_table <- Table$create(airquality)
air_table
```
```{r, test_table_create_from_tibble, opts.label = "test"}
test_that("table_create_from_tibble chunk works as expected", {
  expect_s3_class(air_table, "Table")
})
```

## Converting data from an Arrow Table to a tibble

You may want to convert an Arrow Table to a tibble to view the data or work with it in your usual analytics pipeline.  You can use either `dplyr::collect()` or `as.data.frame()` to do this.

```{r, collect_table}
air_tibble <- dplyr::collect(air_table)
air_tibble
```
```{r, test_collect_table, opts.label = "test"}
test_that("collect_table chunk works as expected", {
  expect_identical(air_tibble, airquality) 
})
```

## Reading and Writing Parquet Files

### Writing a Parquet file

You can write Parquet files to disk using `arrow::write_parquet()`.
```{r, write_parquet}
# Create table
my_table <- Table$create(tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
# Write to Parquet
write_parquet(my_table, "my_table.parquet")
```
```{r, test_write_parquet, opts.label = "test"}
test_that("write_parquet chunk works as expected", {
  expect_true(file.exists("my_table.parquet"))
})
```
 
### Reading a Parquet file

Given a Parquet file, it can be read back in by using `arrow::read_parquet()`.

```{r, read_parquet}
parquet_tbl <- read_parquet("my_table.parquet")
parquet_tbl
```
```{r, test_read_parquet, opts.label = "test"}
test_that("read_parquet works as expected", {
  expect_identical(parquet_tbl, tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
})
```

As the argument `as_data_frame` was left set to its default value of `TRUE`, the file was read in as a `data.frame` object.

```{r, read_parquet_2}
class(parquet_tbl)
```
```{r, test_read_parquet_2, opts.label = "test"}
test_that("read_parquet_2 works as expected", {
  expect_s3_class(parquet_tbl, "data.frame")
})
```
If you set `as_data_frame` to `FALSE`, the file will be read in as an Arrow Table.

```{r, read_parquet_table}
my_table_arrow_table <- read_parquet("my_table.parquet", as_data_frame = FALSE)
my_table_arrow_table
```

```{r, read_parquet_table_class}
class(my_table_arrow_table)
```
```{r, test_read_parquet_table_class, opts.label = "test"}
test_that("read_parquet_table_class works as expected", {
  expect_s3_class(my_table_arrow_table, "Table")
})
```

### How to read a Parquet file from S3 

You can open a Parquet file saved on S3 by calling `read_parquet()` and passing the relevant URI as the `file` argument.

```{r, read_parquet_s3, eval = FALSE}
df <- read_parquet(file = "s3://ursa-labs-taxi-data/2019/06/data.parquet")
```
For more in-depth instructions, including how to work with S3 buckets which require authentication, you can find a guide to reading and writing to/from S3 buckets here: https://arrow.apache.org/docs/r/articles/fs.html.

### How to filter columns while reading a Parquet file 

When reading in a Parquet file, you can specify which columns to read in via the `col_select` argument.

```{r, read_parquet_filter}
# Create table to read back in 
dist_time <- Table$create(tibble::tibble(distance = c(12.2, 15.7, 14.2), time = c(43, 44, 40)))
# Write to Parquet
write_parquet(dist_time, "dist_time.parquet")

# Read in only the "time" column
time_only <- read_parquet("dist_time.parquet", col_select = "time")
time_only
```
```{r, test_read_parquet_filter, opts.label = "test"}
test_that("read_parquet_filter works as expected", {
  expect_identical(time_only, tibble::tibble(time = c(43, 44, 40)))
})
```

## Reading and Writing Feather files 

### Write an IPC/Feather V2 file

The Arrow IPC file format is identical to the Feather version 2 format.  If you call `write_arrow()`, you will get a warning telling you to use `write_feather()` instead.

```{r, write_arrow}
# Create table
my_table <- Table$create(tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
write_arrow(my_table, "my_table.arrow")
```
```{r, test_write_arrow, opts.label = "test"}
test_that("write_arrow chunk works as expected", {
  expect_true(file.exists("my_table.arrow"))
  expect_warning(
    write_arrow(iris, "my_table.arrow"),
    regexp = "Use 'write_ipc_stream' or 'write_feather' instead."
  )
})
```

Instead, you can use `write_feather()`.

```{r, write_feather}
my_table <- Table$create(tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
write_feather(my_table, "my_table.arrow")
```
```{r, test_write_feather, opts.label = "test"}
test_that("write_feather chunk works as expected", {
  expect_true(file.exists("my_table.arrow"))
})
```
### Write a Feather (version 1) file

For legacy support, you can write data in the original Feather format by setting the `version` parameter to `1`.

```{r, write_feather1}
# Create table
my_table <- Table$create(tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
# Write to Feather format V1
write_feather(mtcars, "my_table.feather", version = 1)
```
```{r, test_write_feather1, opts.label = "test"}
test_that("write_feather1 chunk works as expected", {
  expect_true(file.exists("my_table.feather"))
})
```

### Read a Feather file

You can read Feather files in via `read_feather()`.

```{r, read_feather}
my_feather_tbl <- read_feather("my_table.arrow")
```
```{r, test_read_feather, opts.label = "test"}
test_that("read_feather chunk works as expected", {
  expect_identical(dplyr::collect(my_feather_tbl), tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
})
```

## Reading and Writing Streaming IPC Files

You can write to the IPC stream format using `write_ipc_stream()`.

```{r, write_ipc_stream}
# Create table
my_table <- Table$create(tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
# Write to IPC stream format
write_ipc_stream(my_table, "my_table.arrows")
```
```{r, test_write_ipc_stream, opts.label = "test"}
test_that("write_ipc_stream chunk works as expected", {
  expect_true(file.exists("my_table.arrows"))
})
```
You can read from IPC stream format using `read_ipc_stream()`.

```{r, read_ipc_stream}
my_ipc_stream <- arrow::read_ipc_stream("my_table.arrows")
```
```{r, test_read_ipc_stream, opts.label = "test"}
test_that("read_ipc_stream chunk works as expected", {
  expect_equal(my_ipc_stream, tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
})
```

## Reading and Writing CSV files 

You can use `write_csv_arrow()` to save an Arrow Table to disk as a CSV.

```{r, write_csv_arrow}
write_csv_arrow(cars, "cars.csv")
```
```{r, test_write_csv_arrow, opts.label = "test"}
test_that("write_csv_arrow chunk works as expected", {
  expect_true(file.exists("cars.csv"))
})
```

You can use `read_csv_arrow()` to read in a CSV file as an Arrow Table.

```{r, read_csv_arrow}
my_csv <- read_csv_arrow("cars.csv", as_data_frame = FALSE)
```

```{r, test_read_csv_arrow, opts.label = "test"}
test_that("read_csv_arrow chunk works as expected", {
  expect_equivalent(dplyr::collect(my_csv), cars)
})
```

## Reading and Writing Partitioned Data 

### Writing Partitioned Data

You can use `write_dataset()` to save data to disk in partitions based on columns in the data.

```{r, write_dataset}
write_dataset(airquality, "airquality_partitioned", partitioning = c("Month", "Day"))
list.files("airquality_partitioned")
```
```{r, test_write_dataset, opts.label = "test"}
test_that("write_dataset chunk works as expected", {
  # Partition by month
  expect_identical(list.files("airquality_partitioned"), c("Month=5", "Month=6", "Month=7", "Month=8", "Month=9"))
  # We have enough files
  expect_equal(length(list.files("airquality_partitioned", recursive = TRUE)), 153)
})
```
As you can see, this has created folders based on the first partition variable supplied, `Month`.

If you take a look in one of these folders, you will see that the data is then partitioned by the second partition variable, `Day`.

```{r}
list.files("airquality_partitioned/Month=5")
```

Each of these folders contains 1 or more Parquet files containing the relevant partition of the data.

```{r}
list.files("airquality_partitioned/Month=5/Day=10")
```

### Reading Partitioned Data

You can use `open_dataset()` to read partitioned data.

```{r, open_dataset}
# Read data from directory
air_data <- open_dataset("airquality_partitioned")

# View data
air_data
```
```{r, test_open_dataset, opts.label = "test"}
test_that("open_dataset chunk works as expected", {
  expect_equal(nrow(air_data), 153)
  expect_equal(arrange(collect(air_data), Month, Day), arrange(airquality, Month, Day), ignore_attr = TRUE)
})
```

