# Reading and Writing Data

This chapter contains recipes related to reading and writing data using Apache 
Arrow.  When reading files into R using Apache Arrow, you can choose to read in 
your file as either a `tibble` or as an Arrow Table object.

There are a number of circumstances in which you may want to read in the data as an Arrow Table:
* your dataset is large and if you load it into memory, it may lead to performance issues
* you want faster performance from your `dplyr` queries
* you want to be able to take advantage of Arrow's compute functions

## Converting from a tibble to an Arrow Table

You want to convert an existing `tibble` or `data.frame` into an Arrow Table.

### Solution

```{r, table_create_from_tibble}
air_table <- Table$create(airquality)
air_table
```
```{r, test_table_create_from_tibble, opts.label = "test"}
test_that("table_create_from_tibble chunk works as expected", {
  expect_s3_class(air_table, "Table")
})
```

## Converting data from an Arrow Table to a tibble

You want to convert an Arrow Table to a tibble to view the data or work with it
in your usual analytics pipeline.  You can use either `dplyr::collect()` or 
`as.data.frame()` to do this.

### Solution

```{r, collect_table}
air_tibble <- dplyr::collect(air_table)
air_tibble
```
```{r, test_collect_table, opts.label = "test"}
test_that("collect_table chunk works as expected", {
  expect_identical(air_tibble, airquality) 
})
```

## Writing a Parquet file

You want to write Parquet files to disk.

### Solution

```{r, write_parquet}
# Create table
my_table <- Table$create(tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
# Write to Parquet
write_parquet(my_table, "my_table.parquet")
```
```{r, test_write_parquet, opts.label = "test"}
test_that("write_parquet chunk works as expected", {
  expect_true(file.exists("my_table.parquet"))
})
```
 
## Reading a Parquet file

You want to read a Parquet file.

### Solution

```{r, read_parquet}
parquet_tbl <- read_parquet("my_table.parquet")
parquet_tbl
```
```{r, test_read_parquet, opts.label = "test"}
test_that("read_parquet works as expected", {
  expect_identical(parquet_tbl, tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
})
```

As the argument `as_data_frame` was left set to its default value of `TRUE`, the file was read in as a `data.frame` object.

```{r, read_parquet_2}
class(parquet_tbl)
```
```{r, test_read_parquet_2, opts.label = "test"}
test_that("read_parquet_2 works as expected", {
  expect_s3_class(parquet_tbl, "data.frame")
})
```

### Discussion

If you set `as_data_frame` to `FALSE`, the file will be read in as an Arrow Table.

```{r, read_parquet_table}
my_table_arrow_table <- read_parquet("my_table.parquet", as_data_frame = FALSE)
my_table_arrow_table
```

```{r, read_parquet_table_class}
class(my_table_arrow_table)
```
```{r, test_read_parquet_table_class, opts.label = "test"}
test_that("read_parquet_table_class works as expected", {
  expect_s3_class(my_table_arrow_table, "Table")
})
```

## Read a Parquet file from S3 

You want to read a Parquet file from S3.

### Solution

```{r, read_parquet_s3, eval = FALSE}
df <- read_parquet(file = "s3://ursa-labs-taxi-data/2019/06/data.parquet")
```

### See also

For more in-depth instructions, including how to work with S3 buckets which require authentication, you can find a guide to reading and writing to/from S3 buckets here: https://arrow.apache.org/docs/r/articles/fs.html.

## Filter columns while reading a Parquet file 

You want to specify which columns to include when reading in a Parquet file.

### Solution

```{r, read_parquet_filter}
# Create table to read back in 
dist_time <- Table$create(tibble::tibble(distance = c(12.2, 15.7, 14.2), time = c(43, 44, 40)))
# Write to Parquet
write_parquet(dist_time, "dist_time.parquet")

# Read in only the "time" column
time_only <- read_parquet("dist_time.parquet", col_select = "time")
time_only
```
```{r, test_read_parquet_filter, opts.label = "test"}
test_that("read_parquet_filter works as expected", {
  expect_identical(time_only, tibble::tibble(time = c(43, 44, 40)))
})
```

## Write an IPC/Feather V2 file

You want to read in a Feather file.

### Solution

```{r, write_feather}
my_table <- Table$create(tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
write_feather(my_table, "my_table.arrow")
```
```{r, test_write_feather, opts.label = "test"}
test_that("write_feather chunk works as expected", {
  expect_true(file.exists("my_table.arrow"))
})
```
### Discussion

For legacy support, you can write data in the original Feather format by setting the `version` parameter to `1`.

```{r, write_feather1}
# Create table
my_table <- Table$create(tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
# Write to Feather format V1
write_feather(mtcars, "my_table.feather", version = 1)
```
```{r, test_write_feather1, opts.label = "test"}
test_that("write_feather1 chunk works as expected", {
  expect_true(file.exists("my_table.feather"))
})

unlink("my_table.feather")
```

## Read a Feather file

You want to read a Feather file.

### Solution

```{r, read_feather}
my_feather_tbl <- read_feather("my_table.arrow")
```
```{r, test_read_feather, opts.label = "test"}
test_that("read_feather chunk works as expected", {
  expect_identical(dplyr::collect(my_feather_tbl), tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99)))
})
unlink("my_table.arrow")
```

## Write Streaming IPC Files

You want to write to the IPC stream format.

### Solution

```{r, write_ipc_stream}
# Create table
my_table <- Table$create(
  tibble::tibble(
    group = c("A", "B", "C"),
    score = c(99, 97, 99)
    )
)
# Write to IPC stream format
write_ipc_stream(my_table, "my_table.arrows")
```
```{r, test_write_ipc_stream, opts.label = "test"}
test_that("write_ipc_stream chunk works as expected", {
  expect_true(file.exists("my_table.arrows"))
})
```

## Read Streaming IPC Files

You want to read from the IPC stream format.

### Solution
```{r, read_ipc_stream}
my_ipc_stream <- arrow::read_ipc_stream("my_table.arrows")
```
```{r, test_read_ipc_stream, opts.label = "test"}
test_that("read_ipc_stream chunk works as expected", {
  expect_equal(
    my_ipc_stream,
    tibble::tibble(group = c("A", "B", "C"), score = c(99, 97, 99))
  )
})
unlink("my_table.arrows")
```

## Reading and Writing CSV files 

You can use `write_csv_arrow()` to save an Arrow Table to disk as a CSV.

```{r, write_csv_arrow}
write_csv_arrow(cars, "cars.csv")
```
```{r, test_write_csv_arrow, opts.label = "test"}
test_that("write_csv_arrow chunk works as expected", {
  expect_true(file.exists("cars.csv"))
})
```

You can use `read_csv_arrow()` to read in a CSV file as an Arrow Table.

```{r, read_csv_arrow}
my_csv <- read_csv_arrow("cars.csv", as_data_frame = FALSE)
```

```{r, test_read_csv_arrow, opts.label = "test"}
test_that("read_csv_arrow chunk works as expected", {
  expect_equivalent(dplyr::collect(my_csv), cars)
})
unlink("cars.csv")
```

## Read JSON files 

You want to read a JSON file.

### Solution

```{r, read_json_arrow}
# Create a file to read back in 
tf <- tempfile()
writeLines('
    {"country": "United Kingdom", "code": "GB", "long": -3.44, "lat": 55.38}
    {"country": "France", "code": "FR", "long": 2.21, "lat": 46.23}
    {"country": "Germany", "code": "DE", "long": 10.45, "lat": 51.17}
  ', tf, useBytes = TRUE)

# Read in the data
countries <- read_json_arrow(tf, col_select = c("country", "long", "lat"))
countries
```
```{r, test_read_json_arrow, opts.label = "test"}
test_that("read_json_arrow chunk works as expected", {
  expect_equivalent(
    countries,
    tibble::tibble(
      country = c("United Kingdom", "France", "Germany"),
      long = c(-3.44, 2.21, 10.45),
      lat = c(55.38, 46.23, 51.17)
    )
  )
})
unlink(tf)
```


## Write Partitioned Data

You want to save data to disk in partitions based on columns in the data.

### Solution

```{r, write_dataset}
write_dataset(airquality, "airquality_partitioned", partitioning = c("Month", "Day"))
list.files("airquality_partitioned")
```
```{r, test_write_dataset, opts.label = "test"}
test_that("write_dataset chunk works as expected", {
  # Partition by month
  expect_identical(list.files("airquality_partitioned"), c("Month=5", "Month=6", "Month=7", "Month=8", "Month=9"))
  # We have enough files
  expect_equal(length(list.files("airquality_partitioned", recursive = TRUE)), 153)
})
```
As you can see, this has created folders based on the first partition variable supplied, `Month`.

If you take a look in one of these folders, you will see that the data is then partitioned by the second partition variable, `Day`.

```{r}
list.files("airquality_partitioned/Month=5")
```

Each of these folders contains 1 or more Parquet files containing the relevant partition of the data.

```{r}
list.files("airquality_partitioned/Month=5/Day=10")
```

## Reading Partitioned Data

You want to read partitioned data.

### Solution

```{r, open_dataset}
# Read data from directory
air_data <- open_dataset("airquality_partitioned")

# View data
air_data
```
```{r, test_open_dataset, opts.label = "test"}
test_that("open_dataset chunk works as expected", {
  expect_equal(nrow(air_data), 153)
  expect_equal(arrange(collect(air_data), Month, Day), arrange(airquality, Month, Day), ignore_attr = TRUE)
})
```

```{r}
unlink("airquality_partitioned", recursive = TRUE)
```

