<!---
  Licensed to the Apache Software Foundation (ASF) under one
  or more contributor license agreements.  See the NOTICE file
  distributed with this work for additional information
  regarding copyright ownership.  The ASF licenses this file
  to you under the Apache License, Version 2.0 (the
  "License"); you may not use this file except in compliance
  with the License.  You may obtain a copy of the License at

    http://www.apache.org/licenses/LICENSE-2.0

  Unless required by applicable law or agreed to in writing,
  software distributed under the License is distributed on an
  "AS IS" BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
  KIND, either express or implied.  See the License for the
  specific language governing permissions and limitations
  under the License.
-->

# Reading and Writing Data - Single Files

## Introduction

When reading files into R using Apache Arrow, you can read:

* a single file into memory as a data frame or an Arrow Table
* a single file that is too large to fit in memory as an Arrow Dataset
* multiple and partitioned files as an Arrow Dataset

This chapter contains recipes related to using Apache Arrow to read and 
write single file data into memory as an Arrow Table. There are a number of circumstances in
which you may want to read in single file data as an Arrow Table:

* your data file is large and having performance issues
* you want faster performance from your `dplyr` queries
* you want to be able to take advantage of Arrow's compute functions

If a single data file is too large to load into memory, you can use the Arrow Dataset API. 
Recipes for using `open_dataset()` and `write_dataset()` are in the Reading and Writing Data - Multiple Files
chapter.

## Convert data from a data frame to an Arrow Table

You want to convert an existing `data.frame` or `tibble` object into an Arrow Table.

### Solution

```{r, table_create_from_df}
air_table <- arrow_table(airquality)
air_table
```
```{r, test_table_create_from_df, opts.label = "test"}
test_that("table_create_from_df chunk works as expected", {
  expect_s3_class(air_table, "Table")
})
```

## Convert data from an Arrow Table to a data frame

You want to convert an Arrow Table to a data frame to view the data or work with it
in your usual analytics pipeline. 

### Solution

```{r, asdf_table}
air_df <- as.data.frame(air_table)
air_df
```
```{r, test_asdf_table, opts.label = "test"}
test_that("asdf_table chunk works as expected", {
  expect_identical(air_df, airquality)
})
```

### Discussion

You can use either `as.data.frame()` or `dplyr::collect()` to do this.

## Write a Parquet file

You want to write a single Parquet file to disk.

### Solution

```{r, write_parquet}
# Create table
my_table <- arrow_table(data.frame(group = c("A", "B", "C"), score = c(99, 97, 99)))
# Write to Parquet
write_parquet(my_table, "my_table.parquet")
```
```{r, test_write_parquet, opts.label = "test"}
test_that("write_parquet chunk works as expected", {
  expect_true(file.exists("my_table.parquet"))
})
```
 
## Read a Parquet file

You want to read a single Parquet file into memory.

### Solution

```{r, read_parquet}
parquet_tbl <- read_parquet("my_table.parquet")
parquet_tbl
```
```{r, test_read_parquet, opts.label = "test"}
test_that("read_parquet works as expected", {
  expect_identical(parquet_tbl, data.frame(group = c("A", "B", "C"), score = c(99, 97, 99)))
})
```

As the argument `as_data_frame` was left set to its default value of `TRUE`, the file was read in as a `data.frame` object.

```{r, read_parquet_2}
class(parquet_tbl)
```
```{r, test_read_parquet_2, opts.label = "test"}
test_that("read_parquet_2 works as expected", {
  expect_s3_class(parquet_tbl, "data.frame")
})
```

### Discussion

If you set `as_data_frame` to `FALSE`, the file will be read in as an Arrow Table.

```{r, read_parquet_table}
my_table_arrow <- read_parquet("my_table.parquet", as_data_frame = FALSE)
my_table_arrow
```


```{r, read_parquet_table_class}
class(my_table_arrow)
```
```{r, test_read_parquet_table_class, opts.label = "test"}
test_that("read_parquet_table_class works as expected", {
  expect_s3_class(my_table_arrow, "Table")
})
```

## Read a Parquet file from S3 

You want to read a single Parquet file from S3 into memory.

### Solution

```{r, read_parquet_s3, eval = FALSE}
df <- read_parquet(file = "s3://voltrondata-labs-datasets/nyc-taxi/year=2019/month=6/part-0.parquet")
```

### See also

For more in-depth instructions, including how to work with S3 buckets which require authentication, you can find a guide to reading and writing to/from S3 buckets here: https://arrow.apache.org/docs/r/articles/fs.html.

## Filter columns while reading a Parquet file 

You want to specify which columns to include when reading in a single Parquet file into memory.

### Solution

```{r, read_parquet_filter}
# Create table to read back in
dist_time <- arrow_table(data.frame(distance = c(12.2, 15.7, 14.2), time = c(43, 44, 40)))
# Write to Parquet
write_parquet(dist_time, "dist_time.parquet")

# Read in only the "time" column
time_only <- read_parquet("dist_time.parquet", col_select = "time")
time_only
```
```{r, test_read_parquet_filter, opts.label = "test"}
test_that("read_parquet_filter works as expected", {
  expect_identical(time_only, data.frame(time = c(43, 44, 40)))
})
```

## Write a Feather V2/Arrow IPC file

You want to write a single Feather V2 file (also called Arrow IPC file).

### Solution

```{r, write_feather}
my_table <- arrow_table(data.frame(group = c("A", "B", "C"), score = c(99, 97, 99)))
write_feather(my_table, "my_table.arrow")
```
```{r, test_write_feather, opts.label = "test"}
test_that("write_feather chunk works as expected", {
  expect_true(file.exists("my_table.arrow"))
})
```
### Discussion

For legacy support, you can write data in the original Feather format by setting the `version` parameter to `1`.

```{r, write_feather1}
# Create table
my_table <- arrow_table(data.frame(group = c("A", "B", "C"), score = c(99, 97, 99)))
# Write to Feather format V1
write_feather(mtcars, "my_table.feather", version = 1)
```
```{r, test_write_feather1, opts.label = "test"}
test_that("write_feather1 chunk works as expected", {
  expect_true(file.exists("my_table.feather"))
})
```

## Read a Feather/Arrow IPC file

You want to read a single Feather V1 or V2 file into memory (also called Arrow IPC file).

### Solution

```{r, read_feather}
my_feather_tbl <- read_feather("my_table.arrow")
```
```{r, test_read_feather, opts.label = "test"}
test_that("read_feather chunk works as expected", {
  expect_identical(as.data.frame(my_feather_tbl), data.frame(group = c("A", "B", "C"), score = c(99, 97, 99)))
})
unlink("my_table.arrow")
```

## Write streaming Arrow IPC files

You want to write to the Arrow IPC stream format.

### Solution

```{r, write_ipc_stream}
# Create table
my_table <- arrow_table(
  data.frame(
    group = c("A", "B", "C"),
    score = c(99, 97, 99)
    )
)
# Write to IPC stream format
write_ipc_stream(my_table, "my_table.arrows")
```
```{r, test_write_ipc_stream, opts.label = "test"}
test_that("write_ipc_stream chunk works as expected", {
  expect_true(file.exists("my_table.arrows"))
})
```

## Read streaming Arrow IPC files

You want to read from the Arrow IPC stream format.

### Solution
```{r, read_ipc_stream}
my_ipc_stream <- arrow::read_ipc_stream("my_table.arrows")
```
```{r, test_read_ipc_stream, opts.label = "test"}
test_that("read_ipc_stream chunk works as expected", {
  expect_equal(
    my_ipc_stream,
    data.frame(group = c("A", "B", "C"), score = c(99, 97, 99))
  )
})
unlink("my_table.arrows")
```

## Write a CSV file 

You want to write Arrow data to a single CSV file.

### Solution

```{r, write_csv_arrow}
write_csv_arrow(cars, "cars.csv")
```
```{r, test_write_csv_arrow, opts.label = "test"}
test_that("write_csv_arrow chunk works as expected", {
  expect_true(file.exists("cars.csv"))
})
```

## Read a CSV file

You want to read a single CSV file into memory.

### Solution

```{r, read_csv_arrow}
my_csv <- read_csv_arrow("cars.csv", as_data_frame = FALSE)
```

```{r, test_read_csv_arrow, opts.label = "test"}
test_that("read_csv_arrow chunk works as expected", {
  expect_equivalent(as.data.frame(my_csv), cars)
})
unlink("cars.csv")
```

## Read a JSON file

You want to read a JSON file into memory.

### Solution

```{r, read_json_arrow}
# Create a file to read back in
tf <- tempfile()
writeLines('
    {"country": "United Kingdom", "code": "GB", "long": -3.44, "lat": 55.38}
    {"country": "France", "code": "FR", "long": 2.21, "lat": 46.23}
    {"country": "Germany", "code": "DE", "long": 10.45, "lat": 51.17}
  ', tf, useBytes = TRUE)

# Read in the data
countries <- read_json_arrow(tf, col_select = c("country", "long", "lat"))
countries
```
```{r, test_read_json_arrow, opts.label = "test"}
test_that("read_json_arrow chunk works as expected", {
  expect_equivalent(
    countries,
    data.frame(
      country = c("United Kingdom", "France", "Germany"),
      long = c(-3.44, 2.21, 10.45),
      lat = c(55.38, 46.23, 51.17)
    )
  )
})
unlink(tf)
```

## Write a compressed single data file

You want to save a single file, compressed with a specified compression algorithm.

### Solution

```{r, parquet_gzip}
# Create a temporary directory
td <- tempfile()
dir.create(td)

# Write data compressed with the gzip algorithm instead of the default
write_parquet(iris, file.path(td, "iris.parquet"), compression = "gzip")
```

```{r, test_parquet_gzip, opts.label = "test"}
test_that("parquet_gzip", {
  expect_true(file.exists(file.path(td, "iris.parquet")))
})
```

### See also

Some formats write compressed data by default.  For more information 
on the supported compression algorithms and default settings, see:

* `?write_parquet()`
* `?write_feather()`

## Read compressed data

You want to read in a single data file which has been compressed.

### Solution

```{r, read_parquet_compressed}
# Create a temporary directory
td <- tempfile()
dir.create(td)

# Write data which is to be read back in
write_parquet(iris, file.path(td, "iris.parquet"), compression = "gzip")

# Read in data
ds <- read_parquet(file.path(td, "iris.parquet"))
ds
```

```{r, test_read_parquet_compressed, opts.label = "test"}
test_that("read_parquet_compressed", {
  expect_s3_class(ds, "data.frame")
  expect_named(
    ds,
    c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species")
  )
})
```

### Discussion

Note that Arrow automatically detects the compression and you do not have to 
supply it in the call to the `read_*()` or the `open_dataset()` functions.

Although the CSV format does not support compression itself, Arrow supports 
reading in CSV data which has been compressed, if the file extension is `.gz`.

```{r, read_compressed_csv}
# Create a temporary directory
td <- tempfile()
dir.create(td)

# Write data which is to be read back in
write.csv(iris, gzfile(file.path(td, "iris.csv.gz")), row.names = FALSE, quote = FALSE)

# Read in data
ds <- read_csv_arrow(file.path(td, "iris.csv.gz"))
ds
```

```{r, test_read_compressed_csv, opts.label = "test"}
test_that("read_compressed_csv", {
  expect_s3_class(ds, "data.frame")
  expect_named(
    ds,
    c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species")
  )
})
```

```{r cleanup_singlefiles, include = FALSE}
# cleanup
unlink("my_table.arrow")
unlink("my_table.arrows")
unlink("cars.csv")
unlink("my_table.feather")
unlink("my_table.parquet")
unlink("dist_time.parquet")
```